HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-14 paper/main.tex
starting to go over sec 3

@@

\subsection{Annotation Guidelines}\label{sec:guidelines}
In this section we summarise the procedure for annotating machine translation output.

There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
following the
UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
and use native speakers of the target language in order to annotate this output. Since these annotators have to
judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
components (projected UCCA nodes), looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated.
There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
which contains one or more sub-components.

\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes of the tree,
the smallest meaning-bearing units.
Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
The traffic light system makes the marking of the
lexical units as simple as possible.

\begin{compactdesc}
\item[Green] The meaning of the word or phrase has been largely captured.
\item[Orange] the essential meaning has been captured, but some part of the translation is wrong. 
For example, this could often be due to the translated word having the wrong tense, or the wrong morphology. 
\item[Red] The essential meaning of the unit has not been captured.
\end{compactdesc}

\paragraph{Structural nodes} contain other nodes which could be either lexical or structural nodes, 
and they can be labelled as ``Adequate'' or ``Bad''. 
What we are trying to gauge for structural nodes
is if the children of these nodes
relate to each other in the same way in the source sentence and in the translation. 
There are many ways that the relationship between the children might go wrong in translation:
\begin{compactitem}
\item One of the child nodes is missing from the aligned translation. 
\item An extra word or phrase has been inserted into the aligned  translation.
\item The components of the translation are ordered differently 
%to 
from
the source.
\end{compactitem}

@@

\subsection{Comparison with HMEANT}\label{sec:hmeant_comp}

@@

>>>>

\oa{explain how semantic units represent relations}

@@

\subsection{Annotation Procedure}\label{sec:guidelines}

This section summarises the procedure for annotation involved in computing HUME. We mark
the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements for each semantic unit of $s$,
defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
We note that (1) is performed once for every source sentence, irrespective
of the number of its translations we wish to evaluate.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$, induced by the annotation $G$
is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
We note that the translation need not be annotated.

\paragraph{Translation Evaluation.}
The basic strategy for computing HUME is to traverse
the semantic units of the source sentence, and to mark the extent to which
they have been correctly translated. As semantic units correspond to the
semantic relations expressed in the text and the inter-relations between them,
HUME effectively aggregates the judgements of the users into a composite score.

Human annotation of the semantic units requires first deciding whether
a unit is an {\it atomic units}, i.e., that it has no meaning-bearing
sub-units, or a {\it structural unit}. In most cases, atomic units would
correspond to individual words, but they may also correspond to opaque
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units are discarded
from the evaluation.

Atomic units can be labelled as ``Green'' (correct), ``Orange'' (partially correct)
and ``Red'' (incorrect). 
A Green label means that the meaning of the word or phrase has been largely preserved.
An Orange label means that the essential meaning of the unit has been preserved,
but some part of the translation is wrong.
This could often be due to the translated word having the wrong tense,
or the wrong morphology, in a way that impacts little on the understandability of the sentence.
A Red label means that the essential meaning of the unit has not been captured.

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

There are many ways that the relationship between the children might go
wrong in translation:

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
\item One of the sub-units is missing from the translation. For instance, if ``dog bites man''
  is translated into ``dog bites''.
\end{compactitem}

We note that the defined labels are more similar in spirit to adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

\subsection{Annotation Interface.}

As translations may be long, we provide assistance to the annotator.
Where the annotator is asked to judge the translation quality of a unit $u$,
the part of by highlighting the part of the translation which is likely to be
most relevant for the semantic unit in question.
This is done by highlighting the words in the translation which are aligned

using automatically derived word-to-word  in order to project
this annotation to the MT output.

looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated.
There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
which contains one or more sub-components.

In order 
, and use native speakers of the target
language to annotate this output.

Since these annotators have to judge the quality of the translation,
they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comparison with HMEANT}\label{sec:hmeant_comp}

@@

\paragraph{Annotation.}
Source sentences are all in English, and the UCCA annotation was performed by four
computational linguists\oa{were there no other annotators from Prague?}.
